Collocated Snapshot Backup with VM-Centric Deduplication in the Cloud

### 2013-09-26 snapshot_system/fast14_architecture.tex
update

\begin{figure}[htbp]
  \centering
  \epsfig{file=images/deletion.png, width=3in}
  \caption{Approximate deletion}
  \label{fig:deletion_flow}
\end{figure}

@@

We seek a fast solution with low resource usage to delete snapshots and
the VM-centric design simplifies the deletion process. 
Since PDS is small and separated, we can focus on  unreferenced non-PDS chunks within each VM. 
%The PDS data chunks are commonly shared among all VMs and we do not consider them
%during snapshot deletion.  The selection of PDS data chunks is updated periodically independent of snapshot deletion process.
Another resource-saving strategy we propose is
an {\em approximate} deletion strategy to trade deletion accuracy for
speed and resource usage. Our method sacrifices a small percent of storage leakage
to efficiently identify unused chunks.
The algorithm contains three aspects.

\begin{itemize}
\item {\bf Computation for snapshot fingerprint summary.}
Every time there is a new snapshot created,
we compute a Bloom-filter with $z$ bits as the reference summary for all non-PDS chunks used 
in this snapshot. Given $h$ snapshots stored for a VM, there are $h$ summary vectors maintained.

\item {\bf Approximate deletion with fast summary comparison.}
When there is a snapshot deletion,  
we need to identify if  chunks to be deleted from that snapshot
are still used by other snapshots. 
This is done approximately and quickly by comparing the 
reference of deleted chunks  with
the merged Bloom-filter reference summary of other live snapshots.
The merging of live snapshot Bloom-filter bits uses the bitwise OR operator 
and the merged vector still takes $z$ bits.
Since the number of live snapshots $h$ is limited for
each VM, 
the time and memory cost of this comparison is small, linear to the number of chunks to be deleted.

@@

\item {\bf Periodic repair of leakage}.
%[exlpain why second Bloom filter, why scan append store]
Leakage repair is conducted periodically to fix the above approximation error.
This procedure compares the live chunks for each VM with what are truly used in the VM snapshot recipes.
That requires a scanning of all chunks in a VM. Since it is a VM-specific procedure 
the space and time cost is relatively small compared to the traditional mark-sweep procedure~\cite{Guo2011} which scans snapshot 
chunks from all VMs.
For example,
consider each reference consumes 8 bytes plus  1 mark bit, a VM that has 40GB backup data with about
10 million chunks will need less than 85MB of memory to complete a VM-specific mark-sweep process.
\end{itemize}

%{\bf Discussion}
We now estimate the size of storage leakage and how often leak repair needs to be conducted.
Assume that  a VM keeps $h$ snapshots in the backup storage, creates and deletes one snapshot
every day. Let $u$ be the total number of chunks brought by the initial backup for a VM, $\Delta u$ be the average
number of additional unique chunks added from one snapshot to the next snapshot version. Then the total number of unique
chunks used in a VM is about:
\[
U = u + (h-1)\Delta u.
\]

@@

For each snapshot deletion, the number of chunks to be deleted is nearly identical to the number of
newly added chunks $\Delta u$. 
Let $R$ be the total number of runs of approximate deletion between two consecutive 
repairs. We estimate  the total leakage $L$ after $R$ runs as:
\[
L = R \Delta u \epsilon.
\]

@@

For example in our tested dataset,  
each VM keeps $h=10$ snapshots and each snapshot has
about 1-5\% of new data. Thus $\frac{\Delta u}{u} \leq 0.05$. For a 40GB snapshot, $u\approx  10$ million.
Then $U=10.45$ million.
We choose  $\epsilon = 0.01$ and $\tau=0.1$.  From Equation~\ref{eq:falserate}, 
$z=10U=100.45$ million bits. From Equation~\ref{eq:leakrepair}, 
leak repair should be triggered once for every R=290 runs of approximate deletion. 
When one machine hosts 25 VMs and there is one snapshot deletion per day per VM, there would be 
only one full leak repair for one physical machine scheduled for every 12 days. 
Each repair uses at most  90MB memory on average
as discussed earlier and takes a short period of time.

>>>>

\begin{figure}[htbp]
  \centering
  \epsfig{file=images/deletion.pdf, width=3in}
  \caption{Approximate deletion}
  \label{fig:deletion_flow}
\end{figure}

@@

We seek a fast solution with low resource usage to delete snapshots and
our VM-centric design simplifies the deletion process. 
Since PDS is small and separated, we can focus on  unreferenced non-PDS chunks within each VM. 
%The PDS data chunks are commonly shared among all VMs and we do not consider them
%during snapshot deletion.  The selection of PDS data chunks is updated periodically independent of snapshot deletion process.
Another resource-saving strategy we propose is
an {\em approximate} deletion strategy to trade deletion accuracy for
speed and resource usage. Our method sacrifices a small percent of storage leakage
to efficiently identify unused chunks.
The algorithm contains three aspects.

\begin{itemize}
\item {\bf Computation for snapshot reference summary.}
Every time there is a new snapshot created,
we compute a Bloom-filter with $z$ bits as the reference summary vector for all non-PDS chunks used 
in this snapshot,
the items we put into summary vector are all the references that are used in its snapshot and segment recipes.
Foe each VM we preset the vector size according to  estimation of VM image size,
given $h$ snapshots stored for a VM, there are $h$ summary vectors maintained.

\item {\bf Approximate deletion with fast summary comparison.}
When there is a snapshot deletion,  
we need to identify if  chunks to be deleted from that snapshot
are still referenced by other snapshots. 
This is done approximately and quickly by comparing the 
reference of deleted chunks  with
the merged Bloom-filter reference summary of other live snapshots.
The merging of live snapshot Bloom-filter bits uses the bitwise OR operator 
and the merged vector still takes $z$ bits.
Since the number of live snapshots $h$ is limited for
each VM, 
the time and memory cost of this comparison is small, linear to the number of chunks to be deleted.

@@

\item {\bf Periodic repair of leakage}.
%[exlpain why second Bloom filter, why scan append store]
Leakage repair is conducted periodically to fix the above approximation error.
This procedure compares the live chunks for each VM with what are truly used in the VM snapshot recipes.
That requires a scan of all the metadata in a single VM backup. Since it is a VM-specific procedure 
the space and time cost is relatively small compared to the traditional mark-sweep procedure~\cite{Guo2011} which scans snapshot 
chunks from all VMs.
For example,
consider each reference consumes 8 bytes plus  1 mark bit, a VM that has 40GB backup data with about
10 million chunks will need less than 85MB of memory to complete a VM-specific mark-sweep process.
\end{itemize}

%{\bf Discussion}
We now estimate the size of storage leakage and how often leak repair needs to be conducted.
Assume that  a VM keeps $h$ snapshots in the backup storage, creates and deletes one snapshot
every day. Let $u$ be the total number of chunks brought by the initial backup for a VM, $\Delta u$ be the average
number of additional chunks added from one snapshot to the next snapshot version. Then the total number of 
chunks stored in a VM's snapshot store is about:
\[
U = u + (h-1)\Delta u.
\]

@@

For each snapshot deletion, the number of chunks to be deleted is nearly identical to the number of
newly added chunks $\Delta u$. 
Let $R$ be the total number of runs of approximate deletion between two consecutive 
repairs. We estimate  the total leakage $L$ after $R$ runs as:
\[
L = R \epsilon \Delta u.
\]

@@

For example in our tested dataset,  
each VM keeps $h=10$ snapshots and each snapshot adds
about 0.1-4\% of new data. Thus we take ${\Delta u}/{u} \approx 0.02$. For a 40GB snapshot, $u\approx  10$ million.
Then $U=11.8$ million.
We choose  $\epsilon = 0.01$ and $\tau=0.1$.  From Equation~\ref{eq:falserate}, 
each summary vector requires $z=10U=118$ million bits. From Equation~\ref{eq:leakrepair}, 
leak repair should be triggered once for every R=590 runs of approximate deletion. 
When one machine hosts 25 VMs and there is one snapshot deletion per day per VM, there would be 
only one full leak repair for one physical machine scheduled for every 23.6 days. 
Each repair uses at most  90MB memory on average
as discussed earlier and takes a short period of time.

### 2013-09-26 snapshot_system/fast14_evaluation.tex
update

Figure~\ref{fig:exbin-efficiency-graph2} also shows the curve  of VC without local similarity search.
Adding local similarity-guided search increases deduplication efficiency from 93\% to over 96\%.
That is one significant improvement compared to the work in ~\cite{WeiZhangIEEEClound}
which uses the parent segment at the same offset to detect duplicates locally.
The reason is that  
%Without this (i.e. when we only 
when a data  segment is moved to another location on disk, for example when a file is rewritten
rather than modified in place,  dirty-bit based detection would not be able to detect such a movement.
 We have found that in
approximately 1/3 of the VMs in our dataset this movement happens frequently,
so the similarity search becomes important for those VMs.
% In general, our experiments show that
% dirty-bit detection at the segment level can reduce the data size to about 24.14\% of original data, 
% which leads about a 75.86\% reduction.
% Similarity-guided local search can further reduce the data size
% to about 12.05\% of original size, namely it delivers an additional 12.09\% reduction.
% The PDS-based global duplicate detection with $\sigma=2\% $
% can reduce the size further to 8.6\% of original size, namely it 
% delivers an additional 3.45\% reduction.
In general, our experiments show that
dirty-bit detection at the segment level can reduce the data size to about 24.14\% of original data, 
which leads to about a 75.86\% reduction.
Similarity-guided local search can further reduce the data size
to about 12.05\% of original, namely it delivers a 50.08\% reduction to the dirty segments.
The popularity-guided global deduplication with $\sigma=2\% $
can reduce the data further to 8.6\% of its original size, so
it provides additional 28.63\% reduction to the remaining data.

@@

% \begin{table*}[t]
%     \centering
%     \begin{tabular}{c|ccc|ccc}
%     Num. of concurrent      & \multicolumn{3}{c|}{Throughput without}    & \multicolumn{3}{c}{Throughput with} \\
%     backup tasks            & \multicolumn{3}{c|}{I/O throttling (MB/s)} & \multicolumn{3}{c}{I/O throttling (MB/s)} \\ \cline{2-7}
%     per node                & Raw                                        & Snapshot Store & QFS  & Raw                                    & Snapshot Store & QFS  \\ \hline
%     1                       & 1369.6                                     & 148.0          & 18.0 & 171.3                                  & 18.5           & 2.26 \\
%     2                       & 2408.5                                     & 260.2          & 31.7 & 201.3                                  & 21.8           & 2.66 \\
%     4                       & 4101.8                                     & 443.3          & 54.1 & 217.8                                  & 23.5           & 2.87 \\
%     6                       & 5456.5                                     & 589.7          & 72.0 & 224.1                                  & 24.2           & 2.96 \\
%     \end{tabular}
% \caption{Throughput of software layers under different concurrency}
% \label{tab:throughput}
% \end{table*}
\begin{table}[htbp]
\begin{small}
    \centering
    \begin{tabular}{c|ccc}
    Num. of concurrent      & \multicolumn{3}{c}{Throughput without}    \\
    backup tasks            & \multicolumn{3}{c}{I/O throttling (MB/s)} \\ \cline{2-4}
    per machine                & Backup                                     & Snapshot Store & QFS  \\ \hline
    1                       & 1369.6                                     & 148.0          & 35.3 \\
    2                       & 2408.5                                     & 260.2          & 61.7 \\
    4                       & 4101.8                                     & 443.3          & 103.1 \\
    6                       & 5456.5                                     & 589.7          & 143.8 \\
    \end{tabular}
\end{small}
\caption{Throughput of software layers per machine under different concurrency}
\label{tab:throughput}
\end{table}

@@

Figure~\ref{fig:leakage} shows the average accumulated storage leakage in terms of percentage of
storage space per VM caused  by approximate deletions.
The solid line is the predicted leakage using Formula~\ref{eq:leakrepair} from Section~\ref{sect:delete}
while the dashed line is the actual leakage measured during the experiment.
% [OTHER PARAMETERS USED]
After 9 snapshot deletions, the actual leakage reaches 0.01\% and this means that
there is only 1MB space leaked for every 10GB of stored data.

>>>>

Figure~\ref{fig:exbin-efficiency-graph2} also shows the curve  of VC without local similarity search.
Adding local similarity-guided search increases deduplication efficiency from 93\% to over 96\%.
That is one significant improvement compared to the work in ~\cite{WeiZhangIEEE}
which uses the parent segment at the same offset to detect duplicates locally.
The reason is that  
%Without this (i.e. when we only 
when a data  segment is moved to another location on disk, for example when a file is rewritten
rather than modified in place,  dirty-bit based detection would not be able to detect such a movement.
 We have found that in
approximately 1/3 of the VMs in our dataset this movement happens frequently,
so the similarity search becomes important for those VMs.
% In general, our experiments show that
% dirty-bit detection at the segment level can reduce the data size to about 24.14\% of original data, 
% which leads about a 75.86\% reduction.
% Similarity-guided local search can further reduce the data size
% to about 12.05\% of original size, namely it delivers an additional 12.09\% reduction.
% The PDS-based global duplicate detection with $\sigma=2\% $
% can reduce the size further to 8.6\% of original size, namely it 
% delivers an additional 3.45\% reduction.
In general, our experiments show that
dirty-bit detection at the segment level can reduce the data size to about 24.14\% of original data, 
which leads to about a 75.86\% reduction.
Similarity-guided local search can further reduce the data size
to about 12.05\% of original, namely it delivers a 50.08\% reduction to the dirty segments.
The popularity-guided global deduplication with $\sigma=2\% $
can reduce the data further to 8.6\% of its original size, so
it provides additional 28.63\% reduction to the remaining data.

@@

% \begin{table*}[t]
%     \centering
%     \begin{tabular}{c|ccc|ccc}
%     Num. of concurrent      & \multicolumn{3}{c|}{Throughput without}    & \multicolumn{3}{c}{Throughput with} \\
%     backup tasks            & \multicolumn{3}{c|}{I/O throttling (MB/s)} & \multicolumn{3}{c}{I/O throttling (MB/s)} \\ \cline{2-7}
%     per node                & Raw                                        & Snapshot Store & QFS  & Raw                                    & Snapshot Store & QFS  \\ \hline
%     1                       & 1369.6                                     & 148.0          & 18.0 & 171.3                                  & 18.5           & 2.26 \\
%     2                       & 2408.5                                     & 260.2          & 31.7 & 201.3                                  & 21.8           & 2.66 \\
%     4                       & 4101.8                                     & 443.3          & 54.1 & 217.8                                  & 23.5           & 2.87 \\
%     6                       & 5456.5                                     & 589.7          & 72.0 & 224.1                                  & 24.2           & 2.96 \\
%     \end{tabular}
% \caption{Throughput of software layers under different concurrency}
% \label{tab:throughput}
% \end{table*}
\begin{table}[htbp]
\begin{small}
    \centering
    \begin{tabular}{|c|ccc|}
\hline
    Num. of concurrent      & \multicolumn{3}{c|}{Throughput without}    \\
    backup tasks            & \multicolumn{3}{c|}{I/O throttling (MB/s)} \\ \cline{2-4}
    per machine                & Backup                                     & Snapshot Store & QFS  \\ \hline
    1                       & 1369.6                                     & 148.0          & 35.3 \\
    2                       & 2408.5                                     & 260.2          & 61.7 \\
    4                       & 4101.8                                     & 443.3          & 103.1 \\
    6                       & 5456.5                                     & 589.7          & 143.8 \\ \hline
    \end{tabular}
\end{small}
\caption{Throughput of software layers per machine under different concurrency}
\label{tab:throughput}
\end{table}

@@

Figure~\ref{fig:leakage} shows the average accumulated storage leakage in terms of percentage of
storage space per VM caused  by approximate deletions.
The dashed line is the predicted leakage using Formula~\ref{eq:leakrepair} from Section~\ref{sect:delete}
given $\Delta u/u=0.02$,
while the top solid line represents the actual leakage measured during the experiment 
with the same $\Delta u/u=0.02$ setting for summary vectors.
After 9 snapshot deletions, the actual leakage reaches 0.0018\% and this means that
there is only 1.8MB space leaked for every 1GB of stored data. 
We also measure the storage leakage under $\Delta u/u=0.05$ and $0.10$ settings for summary vectors.
As this value increases, larger summary vectors are generated thus results in a decrease on the
false-positive ratio, therefore less amount of unused chunks are misjudged.
